---
title: "Posterior problem with Sigma mixture prior"
author: "Cory Lanker"
date: "4/23/2017"
version: "1.0"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(stats)
library(plotrix)
ellipse.ci <- function(dat, mat, p, col=2, lwd=2){
	g=svd(mat)
	d=atan(g$u[2,1]/g$u[2,2])*180/pi
	q=qnorm(1-(1-p)/2)
	draw.ellipse(dat[2],dat[1],
		q*sqrt(pi/2*g$d[2]),q*sqrt(pi/2*g$d[1]),
		angle=d, border=col, lwd=lwd)		
}

```

## A potential problem with the posterior under some Sigma prior conditions

When there is an overfit input Sigma including in the mixture,
the likelihood and prior could have the same form for that collection
of observations, and there won't be a prior for smoothing the posterior.
This potentially complicates 

1. prediction of new points, since there may be overfit regions
where we would like a smoothed regression function,
2. clustering data points in the training set, as points that
should be in the same group are split due to the likelihood/prior
fixation on the Sigma input prior that matches a subset of the points
of that presumed single cluster.

I'll only consider the impact of duplicate mixture covariance/sample covariance
on prediction (#1 above)

### Files used in these simulation study
Files are found in ~/Dropbox/final/testing
kmixfun_v36.R has the R functions (copied from/based on v35)
Directory 'code' has the C functions
These simulations evolved in files testrun170404.R and testrun170416.R

## Problem Summary

An instability exists in the program when there is an isolated grouping
of observations that has its own input covariance matrix.
The problem occurs when there is a cluster of those points in
the Gibbs sampler with no other points. When this occurs:

1. it is unlikely that other points will join the cluster,
2. it is unlikely that the cluster points leave the cluster,
3. the cluster mean will not change much
4. the cluster covariance will not change at all
5. the cluster proportion is unlikely to become small enough to affect 
any of the above

### Other points stay out of cluster

### Group points stay in cluster

### Cluster mean does not change

### Cluster covariance does not change
```{r covariance}
X1 = mvrnorm(n=10, c(0,0,0), diag(3)*2)
X2 = mvrnorm(n=10, c(0,0,0), diag(3) + .5)
V1 = cov(X1); V2 = cov(X2)
det(V1)
det(V2)
det(V1+V2)
det(V1+V1)
det(V2+V2)
# Difference from Wishart draws
I2 = solve(V2)/7
solve(rWishart(1, 7, I2)[,,1])

S2 = diag(3)
S2[1,3]=S2[3,1]=10; S2[3,3]=150
X2 = mvrnorm(n=10, c(0,0,0), S2)
V2 = cov(X2)
I2 = solve(V2)/7
solve(rWishart(1, 7, I2)[,,1])

```


```{r contours}
f = rWishart(1,10,diag(2)*9)[,,1]
center=c(20,20)
v = mvrnorm(n=1000, mu=center, Sigma=f)
plot.new()
plot(v[,2], v[,1],pch=1,cex=.67,asp=1,
	xlim=range(v[,2])+c(-10,10),ylim=range(v[,1])+c(-10,10))
ellipse.ci(v, f, .95, 'blue', 2)
abline(h=-5:11*10,v=-5:11*10,lty=2)
```
